Uppangala tree functional traits: Article draft

### 2016-07-26 inst/doc/draft.Rmd
v0.1.9023 Mantel results and Ist NM3

---
title: 'Uppangala tree functional traits: Article draft'
author: Sylvain Schmitt sylvain.schmitt@agroparistech.fr
output: pdf_document
bibliography: refs.bib
---

@@

We selected two uncorrelated biotic factors: canopy height and basal area (Spearman's $rho = -0.09$, $p-value > 0.1$). Canopy height was calculated from a digital canopy model built with the 2013 LIDAR campaign on Uppangala plots with resolution $1*1 m^2$. Basal area was calculated from 2013 girth inventory of Uppangala plot on stems with a diameter above 10 centimeters at breast height ($DBH>10~cm$).

@@

### Functional composition of communities

@@

CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative
abundance, $p_{i}$:
$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
CWM was also computed in presence-absence, to get better insight of the role of rare species in community traits distribution. CWV for each trait $t$ was calculated as the difference between the mean of species trait values, $t_{i}$, and the $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
abundance, $p_{i}$:
$$CWV = \sum_{i=1}^{S} p_{i} * (t_{i} - CWM)^2 ~ (2)$$
CVNND was computed in abundance inside each community. For both CWV and CVNND, we calculated the standardized effect size (SES) as: 
$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}}$$
where $I_{obs}$ is the observed metric and $I_{null}$ and $\sigma _{null}$ are the mean and standard deviation of null distributions of NM2 and NM1 for CWV and CVNND respectively.

@@

### Composition variation between communities

@@

$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between communities. Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion. In parrallel variation of functional composition between communities was assessed by $U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007. Observed values were compared to NM1. Negative Hardy's metrics significate that species are functionaly or phylogenetically

@@

# Result tables
kable(regression_table(CWM.lm[[1]], reg, pval[,1], ppval[[1]]))
kable(regression_table(CWM.lm[[2]], reg, pval[,2], ppval[[2]]))
rm(CWM.lm, nCWM.lm, pval, ppval, reg) ; invisible(gc())
```

@@

cl <- makeCluster(3)
clusterExport(cl, list('nL', 'CWM', 'Q'))
nCWM.cwv <- clusterApply(cl, nL, function(l){
  lapply(as.list(names(CWM$abundance)), function(v){
    apply(l, 1, function(x){
      weighted.mean(Q[,v], x, na.rm=T)
    })
  })
})
stopCluster(cl)
nCWM.cwv <- lapply(nCWM.cwv, function(x){names(x) <- names(CWM$abundance) ; return(x)})
nCWV <- mapply(function(l,cwm){
  lapply(names(CWM$abundance), function(v){
    apply(cbind(data.frame(cwm[[which(names(cwm) == v)]]),l), 1, function(x){
      sqrt(weighted.mean((Q[,v]-x[1])^2, x[-1], na.rm=T))
    })
  })
}, l = nL, cwm = nCWM.cwv, SIMPLIFY = F)
nCWV <- lapply(nCWV, function(x){names(x) <- names(CWM$abundance) ; return(x)})
rm(nCWM.cwv, cl, nL) ; invisible(gc())
## Non null CWVs
nCWV <- lapply(nCWV, `[`, names(nCWV[[1]]))
nCWV <- apply(do.call(rbind, nCWV), 2, as.list)
nCWV <- lapply(nCWV, function(x){do.call('cbind', x)})
nCWV <- lapply(nCWV, data.frame)
CWV.pval <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    rank(c(p,q))[1]/(n+1)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = CWV, y = nCWV))
CWV.pval.res <- data.frame(inf = apply(CWV.pval, 2, function(x){length(which(x < 0.025))/length(x)}), sup = apply(CWV.pval, 2, function(x){length(which(x > 0.975))/length(x)}))
## CWV SES
CWV.ses <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    (p - mean(q))/sd(q)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = CWV, y = nCWV))
CWV.ses.lm <- apply(CWV.ses, 2, function(y){lm(y ~., R)})
CWV.ses.lm <- lapply(CWV.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
CWV.ses.res <- lapply(CWV.ses.lm, function(x){summary(x)$coefficients})
CWV.ses.res <- lapply(CWV.ses.res, data.frame)
CWV.ses.res <- lapply(CWV.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
CWV.ses.res <- data.frame(data.table::rbindlist(lapply(CWV.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
CWV.ses.res <- CWV.ses.res[-1]
CWV.ses.res$Slope <- NA
CWV.ses.res <- CWV.ses.res[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
CWV.ses.res <- apply(CWV.ses.res, 2, as.character)
row.names(CWV.ses.res) <- names(CWV.ses)
CWV.ses.res[is.na(CWV.ses.res)] <- ''
CWV.ses.res <- data.frame(CWV.ses.res)
CWV.ses.res$R2 <- round(unlist(lapply(CWV.ses.lm, function(x){summary(x)$r.squared})),3)

# Results
barplot(t(CWV.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
kable(CWV.ses.res)
rm(CWV.pval, CWV.pval.res, CWV.ses.res, CWV.ses.lm, nCWV)
```

@@

# Simulated CVNNDs 
cl <- makeCluster(3)
clusterExport(cl, list('nQ', 'L'))
nCVNND <- clusterApply(cl, nQ, function(x){
  apply(L$abundance, 1, function(l){
    apply(x, 2, function(q){
      cati::CVNND(rep(q,l), na.rm = T)
    })
  })
})
stopCluster(cl) ; rm(cl)
rm(nQ) ; invisible(gc)
## Non null CVNNDs
nCVNND <- lapply(nCVNND, function(x){as.list(data.frame(t(x)))})
nCVNND <- lapply(nCVNND, function(x){lapply(x, function(y){names(y) <- colnames(CVNND) ; return(y)})})
nCVNND <- lapply(nCVNND, `[`, names(nCVNND[[1]]))
nCVNND <- apply(do.call(rbind, nCVNND), 2, as.list)
nCVNND <- lapply(nCVNND, function(x){do.call('cbind', x)})
nCVNND <- lapply(nCVNND, data.frame)
CVNND.pval <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    rank(c(p,q))[1]/(n+1)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
row.names(CVNND.pval) <- colnames(CVNND)
CVNND.pval.res <- data.frame(inf = apply(CVNND.pval, 2, function(x){length(which(x < 0.05))/length(x)}))
## CVNND SES
CVNND.ses <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    (p - mean(q))/sd(q)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
row.names(CVNND.ses) <- colnames(CVNND)
CVNND.ses.lm <- apply(CVNND.ses, 2, function(y){lm(y ~., R)})
CVNND.ses.lm <- lapply(CVNND.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
CVNND.ses.res <- lapply(CVNND.ses.lm, function(x){summary(x)$coefficients})
CVNND.ses.res <- lapply(CVNND.ses.res, data.frame)
CVNND.ses.res <- lapply(CVNND.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
CVNND.ses.res <- data.frame(data.table::rbindlist(lapply(CVNND.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
CVNND.ses.res <- CVNND.ses.res[-1]
CVNND.ses.res$Slope <- NA
CVNND.ses.res <- CVNND.ses.res[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
CVNND.ses.res <- apply(CVNND.ses.res, 2, as.character)
row.names(CVNND.ses.res) <- names(CVNND.ses)
CVNND.ses.res[is.na(CVNND.ses.res)] <- ''
CVNND.ses.res <- data.frame(CVNND.ses.res)
CVNND.ses.res$R2 <- round(unlist(lapply(CVNND.ses.lm, function(x){summary(x)$r.squared})),3)

# Results
barplot(t(CVNND.pval.res), las = 2, col = c('darkblue'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior'), fill = c('darkblue'))
kable(CVNND.ses.res)
rm(CVNND.pval, CVNND.pval.res, CVNND.ses.res, CVNND.ses.lm, nCVNND) ; invisible(gc())
```

@@

**Artefacts of phylogenetic tree building to be removed**

```{r PhyloMaker}
library(phylosignal)
library(phytools)
library(phylobase)
path <- system.file('phylogeny', package = 'Uppangala')
source(file.path(path, 'R_codes for function S.PhyloMaker.txt'))
phylo <- read.tree(file.path(path, 'PhytoPhylo.tre'))
nodes <- read.csv(file.path(path, 'nodes.csv'), h = T)
sp <- Species[which(Species$LatinName %in% row.names(PFT_sp)), c('LatinName', 'Genus', 'Family')]
names(sp) <- c('species', 'genus', 'family')
results <- invisible(S.PhyloMaker(sp, phylo, nodes)) # Humboldtia_brunonis unmatched
tree <- results$Scenario.1
rm(phylor, results, modes, sp, S.PhyloMaker, path)
```

@@

```{r Phylosignal}
data <- PFT_sp
row.names(data) <- gsub(' ', '_', row.names(data), fixed = T)
tips <- row.names(data)[which(is.na(data$WD))]
phy <- drop.tip(tree, tips)
data <- data[phy$tip.label,c(2:5,7)]
p4d <- phylo4d(phy,data)
phyloS <- phyloSignal(p4d, methods = 'I') 
phyloS <- cbind(phyloS$stat, phyloS$pvalue)
names(phyloS) <- c('Moran\'s I', 'p-value')
phyloS$` ` <- lapply(as.list(phyloS$`p-value`), stars, ns = ' ')
kable(phyloS)
rm(data, tips, p4d, phyloS)
```

Variations of both functional and phylogenetic compositions between communities don't significantly differ from null models ($U_{st} = -1.1*10^-3$ with $p-value = 0.114$, $\tau _{st} = 1.8*10^-3$ with $p-value = 0.079$, $\beta _{st} = -8.3*10^-4$ with $p-value = NaN$, and $\pi _{st} = -1.8*10^-3$ with $p-value = 0.247$).

**Following tables will probably be removed in final report by simply including results in text. They are just here for discussion.**

Table 8: Summary of variation of functional composition between communities. _$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of functional composition between communities. Observed values were compared to the one obtained with the null model *"1s"*" suggested by @hardy_testing_2008. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

```{r Functionnal spatial structure, eval=FALSE, include=FALSE}
library(spacodiR)
mtest <- dist(as(quadrats(), 'SpatialPoints')@coords)

# Observed Hardys
spafun <- lapply(as.list(1:length(Q)), function(i){
  if (length(which(is.na(Q[i]))) > 0)
    q.names <- row.names(Q[-which(is.na(Q[i])),])
  else
    q.names <- row.names(Q)
  sel <- intersect(names(L$abundance), q.names)
  return(spacodi.calc(sp.plot = t(L$abundance[sel]), sp.traits = data.frame(Q[sel,i], row.names = sel), pairwise = T))
})
names(spafun) <- names(Q)
spafun <- lapply(spafun, function(x){x[c('pairwise.Ist','pairwise.Ust', 'pairwise.TAUst')]})
## Mantel values
spafun.mantel <- lapply(spafun, function(t){lapply(t, function(m){vegan::mantel(m, mtest, permutations = 0)$statistic})})

# Simulated Hardys
cl <- makeCluster(3)
clusterExport(cl, list('nQ', 'L', 'spacodi.calc'))
nspafun <- clusterApply(cl, nQ, function(q){
  lapply(as.list(1:length(q)), function(i){
    if (length(which(is.na(q[i]))) > 0)
      q.names <- row.names(q[-which(is.na(q[i])),])
    else
      q.names <- row.names(q)
    sel <- intersect(names(L$abundance), q.names)
    return(spacodi.calc(sp.plot = t(L$abundance[sel]), sp.traits = data.frame(q[sel,i], row.names = sel), pairwise = T))
  })
})
stopCluster()
rm(cl, nQ) ; invisible(gc())
## Simulated mantels

# Result
kable(spafun.res)
```

Table 9: Summary of variation of phylogenetic composition between communities. _$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between communities. Observed values were compared to a null model randomizing traits among species (permuting raw in `Trait*Species` matrix). The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

```{r Phylogenetic spatial structure, eval=FALSE, include=FALSE}
for(i in 1:2) phy[[i]]$tip.label <- gsub('_', ' ', phy[[i]]$tip.label, fixed = T)
for(i in 1:2) phy[[i]]$tip.label <- as.character(l$Sp_Code[match(phy[[i]]$tip.label, row.names(l))])
Lm <- lapply(phy, function(x){Lm[intersect(names(Lm), x$tip.label)]})
phyRD <- lapply(phy, function(x){x$node.label <- NULL ; return(x)})
phyRD <- lapply(phyRD, function(x){rep(list(x), 999)})
phyRD <- lapply(phyRD, function(x){lapply(x, resamp.phy)})
LmRD <- lapply(Lm, function(x){rep(list(x), 999)})
LmRD <- mapply(function(x,y){mapply(function(r,s){r[intersect(names(r), s$tip.label)]}, r = x, s = y, SIMPLIFY = F)}, x = LmRD, y = phyRD, SIMPLIFY = F)
phyRD <- mapply(function(x,y){mapply(function(r,s){drop.tip(r, setdiff(r$tip.label, names(s)))}, r = x, s= y, SIMPLIFY = F)}, x = phyRD, y = LmRD, SIMPLIFY = F)

spaphy <- mapply(function(x,y){spacodi.calc(sp.plot = t(y), phy = x)}, x = phy, y = Lm.l, SIMPLIFY = F)
spaphy <- do.call('rbind.data.frame', lapply(spaphy, function(x){x[1:4]}))

spaphy.l <- mapply(function(x,y){mapply(function(r,s){spacodi.calc(t(r), s)}, r = x, s = y, SIMPLIFY = F)},x = LmRD, y = phyRD, SIMPLIFY = F)
spaphy.l <- lapply(spaphy.l, function(x){lapply(x, function(y){y[1:4]})})
spaphyRD <- list(Ist = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Ist}))}),
                 Tst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Tst}))}),
                 Bst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$Bst}))}),
                 PIst = lapply(spaphy.l, function(x){unlist(lapply(x, function(y){y$PIst}))}))
pvalues <- mapply(function(obs,rd){
  mapply(function(o,r){rank(c(o, r))[1]/1000}, o = obs, r = rd)
  }, obs = lapply(as.list(spaphy), as.list), rd = spaphyRD)
row.names(pvalues) <- row.names(spaphy)

pvalues <- as.data.frame(pvalues)[3:4]
pvalues$Bst <-  rep(NaN,2)
spaphy.res <- format(spaphy[3:4], scientific = T, digits = 2)
pvalues <- data.frame(apply(pvalues, c(1,2), function(y){paste(y, stars(y, ns = ' '))}))
spaphy.res$Bst <- paste(spaphy.res$Bst, '/', pvalues$Bst)
spaphy.res$PIst <- paste(spaphy.res$PIst, '/', pvalues$PIst)
kable(spaphy.res)
```

>>>>

---
title: 'Uppangala tree functional traits: Article draft'
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
output: word_document
bibliography: refs.bib
---

@@

We selected two biotic factors: canopy height and basal area. Canopy height was calculated from a digital canopy model built with the 2013 LIDAR campaign on Uppangala plots with resolution $1*1 m^2$. Basal area was calculated from 2013 girth inventory of Uppangala plot on stems with a diameter above 10 centimeters at breast height ($DBH>10~cm$).

@@

### Local functionnal composition of communities

@@

CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative
abundance, $p_{i}$:
$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
CWM was also computed in presence-absence, to get better insight of the role of rare species in community traits distribution. CWV for each trait $t$ was calculated as the difference between the mean of species trait values, $t_{i}$, and the $CWM$, of the $S$ species in each community, with each species, $i$, weighted by its relative
abundance, $p_{i}$:
$$CWV = \sum_{i=1}^{S} p_{i} * (t_{i} - CWM)^2 ~ (2)$$
CVNND for each trait $t$ was calculated as nearest neigbor distances (NND) variance, $\sigma _{NND}$, of the $i$ individuals in each community, divided by its mean, $\mu _{NND}$; and NND for one individual $i$ is calculated as the minimal distance $d$ between the $j$ other individuals of the community :
$$CVNND = \frac{\sigma _{NDD}}{\mu _{NDD}} ~ with ~ NND_{i} = min(d_{i,j})_{i \neq j} ~ (3)$$
For both CWV and CVNND, we calculated the standardized effect size (SES) as: 
$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (4)$$
where $I_{obs}$ is the observed metric and $I_{null}$ and $\sigma _{null}$ are the mean and standard deviation of null distributions of NM2 and NM1 for CWV and CVNND respectively.

@@

### Functionnal and phylogenetic composition between communities

@@

$\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of phylogenetic composition between community pairs. In parrallel variation of functional composition between community pairs was assessed by $U st$ and $\tau st$ metric. Environmental distances between community pairs was also calculated as distances between communities on the plan composed by the two first axes of the principal component analysis of environmental variables (see suppplementary materials S1). Finally spatial distances between community pairs was also assessed. We used Mantel test to sudy the link between functionnal and phylogenetic pairwise compositions with hardy's metrics and environmental distances between community pairs. A second mantel test was realized to assess those pairwise links between environment and functionnal and phylogenetic structure knowing spatial structure with the spatial distances between community pairs matrix.

Observed mantel correlations for phylogenetic composition were compared to 999 obtained with the null model *"1s"*" suggested by @hardy_testing_2008 randomizing phylogenetic tree tips. Observed mantel correlations for functionnal composition were compared to NM1, randomizing traits among species.

@@

# Result tables
kable(bold(regression_table(CWM.lm[[1]], reg, pval[,1], ppval[[1]])))
kable(bold(regression_table(CWM.lm[[2]], reg, pval[,2], ppval[[2]])))
rm(CWM.lm, nCWM.lm, pval, ppval, reg) ; invisible(gc())
```

@@

cl <- makeCluster(3)
clusterExport(cl, list('nL', 'CWM', 'Q'))
nCWM.cwv <- clusterApply(cl, nL, function(l){
  lapply(as.list(names(CWM$abundance)), function(v){
    apply(l, 1, function(x){
      weighted.mean(Q[,v], x, na.rm=T)
    })
  })
})
stopCluster(cl)
nCWM.cwv <- lapply(nCWM.cwv, function(x){names(x) <- names(CWM$abundance) ; return(x)})
nCWV <- mapply(function(l,cwm){
  lapply(names(CWM$abundance), function(v){
    apply(cbind(data.frame(cwm[[which(names(cwm) == v)]]),l), 1, function(x){
      sqrt(weighted.mean((Q[,v]-x[1])^2, x[-1], na.rm=T))
    })
  })
}, l = nL, cwm = nCWM.cwv, SIMPLIFY = F)
nCWV <- lapply(nCWV, function(x){names(x) <- names(CWM$abundance) ; return(x)})
rm(nCWM.cwv, cl, nL) ; invisible(gc())
## Non null CWVs
nCWV <- lapply(nCWV, `[`, names(nCWV[[1]]))
nCWV <- apply(do.call(rbind, nCWV), 2, as.list)
nCWV <- lapply(nCWV, function(x){do.call('cbind', x)})
nCWV <- lapply(nCWV, data.frame)
CWV.pval <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    rank(c(p,q))[1]/(n+1)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = CWV, y = nCWV))
CWV.pval.res <- data.frame(inf = apply(CWV.pval, 2, function(x){length(which(x < 0.025))/length(x)}), sup = apply(CWV.pval, 2, function(x){length(which(x > 0.975))/length(x)}))
## CWV SES
CWV.ses <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    (p - mean(q))/sd(q)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = CWV, y = nCWV))
CWV.ses.lm <- apply(CWV.ses, 2, function(y){lm(y ~., R)})
CWV.ses.lm <- lapply(CWV.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
CWV.ses.res <- lapply(CWV.ses.lm, function(x){summary(x)$coefficients})
CWV.ses.res <- lapply(CWV.ses.res, data.frame)
CWV.ses.res <- lapply(CWV.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
CWV.ses.res <- data.frame(data.table::rbindlist(lapply(CWV.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
CWV.ses.res <- CWV.ses.res[-1]
CWV.ses.res$Slope <- NA
CWV.ses.res <- CWV.ses.res[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
CWV.ses.res <- apply(CWV.ses.res, 2, as.character)
row.names(CWV.ses.res) <- names(CWV.ses)
CWV.ses.res[is.na(CWV.ses.res)] <- ''
CWV.ses.res <- data.frame(CWV.ses.res)
CWV.ses.res$`$R^2$` <- round(unlist(lapply(CWV.ses.lm, function(x){summary(x)$r.squared})),3)

# Results
barplot(t(CWV.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
kable(bold(CWV.ses.res))
rm(CWV.pval, CWV.pval.res, CWV.ses.res, CWV.ses.lm, nCWV)
```

@@

# Simulated CVNNDs 
cl <- makeCluster(3)
clusterExport(cl, list('nQ', 'L'))
nCVNND <- clusterApply(cl, nQ, function(x){
  apply(L$abundance, 1, function(l){
    apply(x, 2, function(q){
      cati::CVNND(rep(q,l), na.rm = T)
    })
  })
})
stopCluster(cl) ; rm(cl)
rm(nQ) ; invisible(gc)
## Non null CVNNDs
nCVNND <- lapply(nCVNND, function(x){as.list(data.frame(t(x)))})
nCVNND <- lapply(nCVNND, function(x){lapply(x, function(y){names(y) <- colnames(CVNND) ; return(y)})})
nCVNND <- lapply(nCVNND, `[`, names(nCVNND[[1]]))
nCVNND <- apply(do.call(rbind, nCVNND), 2, as.list)
nCVNND <- lapply(nCVNND, function(x){do.call('cbind', x)})
nCVNND <- lapply(nCVNND, data.frame)
CVNND.pval <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    rank(c(p,q))[1]/(n+1)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
row.names(CVNND.pval) <- colnames(CVNND)
CVNND.pval.res <- data.frame(inf = apply(CVNND.pval, 2, function(x){length(which(x < 0.05))/length(x)}))
## CVNND SES
CVNND.ses <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    (p - mean(q))/sd(q)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = as.list(data.frame(t(CVNND))), y = nCVNND))
row.names(CVNND.ses) <- colnames(CVNND)
CVNND.ses.lm <- apply(CVNND.ses, 2, function(y){lm(y ~., R)})
CVNND.ses.lm <- lapply(CVNND.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
CVNND.ses.res <- lapply(CVNND.ses.lm, function(x){summary(x)$coefficients})
CVNND.ses.res <- lapply(CVNND.ses.res, data.frame)
CVNND.ses.res <- lapply(CVNND.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
CVNND.ses.res <- data.frame(data.table::rbindlist(lapply(CVNND.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
CVNND.ses.res <- CVNND.ses.res[-1]
CVNND.ses.res$Slope <- NA
CVNND.ses.res <- CVNND.ses.res[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
CVNND.ses.res <- apply(CVNND.ses.res, 2, as.character)
row.names(CVNND.ses.res) <- names(CVNND.ses)
CVNND.ses.res[is.na(CVNND.ses.res)] <- ''
CVNND.ses.res <- data.frame(CVNND.ses.res)
CVNND.ses.res$`$R^2$` <- round(unlist(lapply(CVNND.ses.lm, function(x){summary(x)$r.squared})),3)

# Results
barplot(t(CVNND.pval.res), las = 2, col = c('darkblue'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior'), fill = c('darkblue'))
kable(bold(CVNND.ses.res))
rm(CVNND.pval, CVNND.pval.res, CVNND.ses.res, CVNND.ses.lm, nCVNND) ; invisible(gc())
```

@@

```{r PhyloMaker}
library(phylosignal)
library(phytools)
library(phylobase)
path <- system.file('phylogeny', package = 'Uppangala')
source(file.path(path, 'R_codes for function S.PhyloMaker.txt'))
phylo <- read.tree(file.path(path, 'PhytoPhylo.tre'))
nodes <- read.csv(file.path(path, 'nodes.csv'), h = T)
sp <- Species[which(Species$LatinName %in% row.names(PFT_sp)), c('LatinName', 'Genus', 'Family')]
names(sp) <- c('species', 'genus', 'family')
results <- invisible(S.PhyloMaker(sp, phylo, nodes)) # Humboldtia_brunonis unmatched
tree <- results$Scenario.1
rm(phylo, results, modes, sp, S.PhyloMaker, path)
```

@@

```{r Phylosignal}
data <- PFT_sp
row.names(data) <- gsub(' ', '_', row.names(data), fixed = T)
tips <- row.names(data)[which(is.na(data$WD))]
phy <- drop.tip(tree, tips)
data <- data[phy$tip.label,c(2:5,7)]
p4d <- phylo4d(phy,data)
phyloS <- phyloSignal(p4d, methods = 'I') 
phyloS <- data.frame(paste(round(phyloS$stat[,1], 3), 
                           unlist(lapply(as.list(phyloS$pvalue[,1]), stars, ns = ' '))),
                     row.names = row.names(phyloS$stat))
names(phyloS) <- 'Moran\'s I'
kable(bold(phyloS))
rm(data, tips, p4d, phyloS, phy)
```

Table 8: Summary of Mantel test between floristic composition, environment and spatial distances between communities. _$Ist$  was used to investigate variation of floristic composition between communities. Observed values were compared to a null model randomizing individuals among subplots while preserving the regional abundance of species, and the number of individuals per community. E is the mantel test of the metric with the environmental distance matrix issued from PCA two first axes. S|E is the mantel partial test of the spatial distance matrix knowing the environment matrix. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

```{r Ist}
library(spacodiR)

# Environmental matrices
XY <- data.frame(as(quadrats(), 'SpatialPoints')@coords)
row.names(XY) <- paste0('C',row.names(XY))
XY <- XY[row.names(L$abundance),]
Ms <- dist(XY)
XY <- pca.env$scores[,1:2]
row.names(XY) <- paste0('C',row.names(XY))
XY <- XY[row.names(L$abundance),]
Me <- dist(XY)
rm(XY)

# Observed Ist
spaflo <- spacodi.calc(sp.plot = t(L$abundance), pairwise = T)$'pairwise.Ist'
## Mantel values
spaflo.mantel <- list(`E` = vegan::mantel(spaflo, Me, permutations = 0)$statistic, 
                      `S | E`= vegan::mantel.partial(spaflo, Ms, Me, permutations = 0)$statistic)

# Simulated mantel
nL <- vegan::permatfull(L$abundance, fixedmar = 'both', mtype = 'count', times = n, shuffle = 'both')
cl <- makeCluster(3)
clusterExport(cl, list('nL', 'spacodi.calc', 'Me', 'Ms'))
nspaflo <- lapply(nL$perm, function(l){spacodi.calc(sp.plot = t(l), pairwise = T)$'pairwise.Ist'})
nspaflo.mantel <- lapply(nspaflo, function(ns){
  list(`E` = vegan::mantel(ns, Me, permutations = 0)$statistic, 
       `S | E`= vegan::mantel.partial(ns, Ms, Me, permutations = 0)$statistic)})
stopCluster(cl)
rm(cl, nL) ; invisible(gc())
## Non null mantels
nspaflo.mantel <- lapply(nspaflo.mantel, `[`, names(nspaflo.mantel[[1]]))
nspaflo.mantel <- apply(do.call(rbind, nspaflo.mantel), 2, as.list)
nspaflo.mantel <- lapply(nspaflo.mantel, unlist)
spaflo.res <- mapply(function(s, ns){
    paste(format(s, scientific = T, digits = 3),
          Uppangala::stars(min(1-rank(c(s,ns))[1]/(n+1),rank(c(s,ns))[1]/(n+1)), ns = '   '))
}, s = spaflo.mantel, ns = nspaflo.mantel)
spaflo.res <- data.frame(spaflo.res)
names(spaflo.res) <- c('$I_{st}$')

# Result
kable(bold(spaflo.res))
rm(mtest, spaflo.mantel, nspaflo, nspaflo.mantel, spaflo.res)
```

Table 9: Summary of Mantel test between functionnal composition, environment and spatial distances between communities. _$U st$ and $\tau st$ metric proposed by @hardy_characterizing_2007 were used to investigate variation of functional composition between communities. Observed values were compared to the null model NM1 randomizing traits among species (permuting raw in `Trait*Species` matrix). E is the mantel test of the metric with the environmental distance matrix issued from PCA two first axes. S|E is the mantel partial test of the spatial distance matrix knowing the environment matrix. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

```{r Ust & TAUst}
# Observed Hardys
spafun <- lapply(as.list(1:length(Q)), function(i){
  if (length(which(is.na(Q[i]))) > 0) {
    sel <- intersect(names(L$abundance), row.names(Q[-which(is.na(Q[i])),]))
    return(spacodi.calc(sp.plot = t(L$abundance[sel]), sp.traits = data.frame(Q[sel,i], row.names = sel), pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
  } else {
    return(spacodi.calc(sp.plot = t(L$abundance), sp.traits = Q, pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
  }
})
names(spafun) <- names(Q)
## Mantel values
spafun.mantel <- lapply(spafun, function(t){list(`E` = lapply(t, function(M){vegan::mantel(M, Me, permutations = 0)$statistic}), `S | E` = lapply(t, function(M){vegan::mantel.partial(M, Ms, Me, permutations = 0)$statistic}))})

# Simulated mantel
cl <- makeCluster(3)
clusterExport(cl, list('nQ', 'L', 'Me', 'Ms', 'spacodi.calc'))
nspafun.mantel <- clusterApply(cl, nQ, function(q){
  s <- lapply(as.list(1:length(q)), function(i){
    if (length(which(is.na(q[i]))) > 0) {
      sel <- intersect(names(L$abundance), row.names(q[-which(is.na(q[i])),]))
      return(spacodi.calc(sp.plot = t(L$abundance[sel]), sp.traits = data.frame(q[sel,i], row.names = sel), pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
    } else {
      return(spacodi.calc(sp.plot = t(L$abundance), sp.traits = q, pairwise = T)[c('pairwise.Ust', 'pairwise.TAUst')])
    }
  })
  names(s) <- names(q)
  res <- list(
    `E` = lapply(s, function(t){
      lapply(t, function(M){
        try(vegan::mantel(M, Me, permutations = 0)$statistic)
        })
      }),
    `S | E` = lapply(s, function(t){
      lapply(t, function(M){
        try(vegan::mantel.partial(M, Ms, Me, permutations = 0)$statistic)
        })
      })
    )
  res <- lapply(res, function(r){names(r) <- names(q) ; return(r)})
  res <- lapply(res, `[`, names(res[[1]]))
  res <- apply(do.call(rbind, res), 2, as.list)
  return(res)
})
stopCluster(cl)
rm(cl, nQ) ; invisible(gc())
